---
title: "Assignment3"
author: "yutian"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r library, message=FALSE}
library(stringr)
library(rvest)
library(httr)
library(tidyverse)
library(xml2)
```

# API

```{r Counts}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")

# Turning it into text
counts <- as.character(counts)
```

## Q1: How many papers were you able to find?
```{r extract data}
# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```

There are 3997 papers under the term “sars-cov-2 trial vaccine.

## Q2: Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.


```{r query_id}
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db     = "pubmed", 
    term   = "sars-cov-2 trial vaccine", 
    retmax = 250
    )
)
# Extracting the content of the response of GET
ids <- httr::content(query_ids)
```

```{r get id details}
ids <- as.character(ids)
ids <- stringr::str_extract_all(ids, "<Id>[0-9]+</Id>")[[1]]
ids <- stringr::str_remove_all(ids, "<Id>|</Id>")
```

```{r publication}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db      = "pubmed",
    id      = paste(ids, collapse = ","),
    rettype = "abstract"
    )
)
publications <- httr::content(publications)
publications_txt <- as.character(publications)

```

```{r Details}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)

```

```{r titles}
titles <- str_extract(pub_char_list, "<ArticleTitle>(\\n|.)+</ArticleTitle>")
titles <- str_remove_all(titles, "</?[[:alnum:]]+>")
titles <- str_replace_all(titles, "\\s+"," ")
```



```{r abstracts}
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>")
abstracts <- str_replace_all(abstracts, "[[:space:]]+", " ")

```

```{r journals}
journals <- str_extract(pub_char_list, "<Title>(\\n|.)+</Title>")
journals<- str_remove_all(journals, "</?[[:alnum:]]+>")
journals<- str_replace_all(journals, "\\s+"," ")
```

```{r dates}
dates <- str_remove_all(pub_char_list, "\\n") %>%
    str_extract("<PubDate>.*</PubDate>")
dates <-str_remove_all(dates, "</?[[:alnum:]]+>")
#remove new line and spaces
dates <- str_replace_all(dates, "\\s+"," ")
```

Finally, the database

```{r database}
database <- data.frame(
  PubMedId = ids,
  Title    = titles,
  Journal = journals,
  Abstract = abstracts,
  Date = dates
)
knitr::kable(database[1:2,], caption = "Some papers about sars-cov-2 trial vaccine")
```

# Text Mining

A new dataset has been added to the data science data repository https://github.com/USCbiostats/data-science-data/tree/master/03_pubmed. The dataset contains 3241 abstracts from articles across 5 search terms. Your job is to analyse these abstracts to find interesting insights.



```{r library text minig}
library(tidytext)
library(dplyr)
library(ggplot2)
```

```{r pubmed}
if (!file.exists("pubmed.csv")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv", "pubmed.csv", method="libcurl", timeout = 60)
}
pmb <- read.csv("pubmed.csv")
pmb <- as_tibble(pmb)
```
## 1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

```{r top 5 tokens with stop words}
pmb %>%
  unnest_tokens(token, abstract) %>%
  count(token, sort = TRUE) %>%
  top_n(5, n) %>%
  ggplot(aes(n, fct_reorder(token, n))) +
  labs(x = "Frequecy", y = "Top 5 tokens",title= "Top 5 tokens with stop words") +
  geom_col()
```
ALL 5 most frequent words are stop words, so remove them to see the difference 

```{r Top 5 tokens without stop words}
pmb %>%
  unnest_tokens(token, abstract) %>%
  anti_join(stop_words, by = c("token" = "word")) %>%
  count(token, sort = TRUE) %>%
  filter(!grepl(pattern = "^[0-9]+$", x = token)) %>%
  top_n(5, n) %>%
  ggplot(aes(n, fct_reorder(token, n))) +
  labs(x = "Frequency", y = "Top 5 tokens without stop words") +
  geom_col()
```

As a result, there is not any stop words and the Top5 most frequent words are covid, patients, cancer , prostate and disease.

## 2. Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.


```{r top 10 bigram}
pmb %>%
    unnest_ngrams(bigram, abstract, n = 2) %>%
    separate(bigram, into = c("first", "second"), sep = " ",remove = FALSE) %>%
    anti_join(stop_words, by = c(first = "word")) %>%
    anti_join(stop_words, by = c(second = "word")) %>%
   filter(str_detect(first, "[a-z]") &
         str_detect(second, "[a-z]"))%>%
    unite(bigram, c("first", "second"), sep = " ") %>%
    count(bigram, sort = TRUE) %>%
    top_n(10, n) %>%
    ggplot(aes(x = n, y = fct_reorder(bigram, n))) + geom_col() +
    labs(y = "bigram", title = "Top 10 Most Frequent Bigrams without stopwords") 
```


As before, remove the stop words, there are top 10 most frequent bigrams

## 3. Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r tf-idf}
pmb %>%
    unnest_tokens(token, abstract) %>%
    anti_join(stop_words, by = c(token = "word")) %>%
    filter(!grepl(pattern = "^[0-9]+$", x = token)) %>%
    group_by(term) %>%
    count(token, sort = TRUE) %>%
  bind_tf_idf(token, term, n) %>%
  arrange(term, desc(tf_idf)) %>%
  top_n(5, tf_idf) %>%
knitr::kable()

```

IN question 1, the top 5 common words are covid, patients, cancer, prostate and disease
After giving the weigh to the words, the tokens are covid, pandemic, coronavirus, sars and cov which  means word "patient","prostate", and "disease" are been replaced. It seems like making more sense since them are more common words in all bio medical fields instead of this specific filed.    