---
title: "lab 04"
author: "yutian"
date: "9/7/2022"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r - install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)
```

## Step1. Read in the data

First download and then read in with data.table:fread()

```{r}
if(!file.exists('../lab03/met_all.gz'))
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
met <- data.table::fread("../lab03/met_all.gz")

```


##  Step2. Prep Data

Remove temperatures less than -17C and change elev 9999 to missing value code.

```{r remove-low-temp}
met <- met[temp > -17][elev == 9999.0, elev := NA]
```

Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).

```{r create-ymd, cache = TRUE}
met <- met[ ,  ymd := as.Date(paste(year, month, day, sep = "-"))]
```

Using the data.table::week function, keep the observations of the first week of the month.

```{r}
met[, table(week(ymd))]
met <- met[ week(ymd) == 31 ]
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.

### check no 9999s in other important variables

```{r check-max}
met[, .(
  temp      = max(temp,na.rm=T),
  rh        = max(rh,na.rm=T), 
  wind.sp   = max(wind.sp,na.rm=T), 
  vis.dist  = max(vis.dist,na.rm=T), 
  dew.point = max(dew.point,na.rm=T), 
  lat       = max(lat,na.rm=T), 
  lon       = max(lon,na.rm=T), 
  elev      = max(elev,na.rm=T)
)]
```
There is no 9999s in the dataset



```{r check-avg}
met_avg <- met[, .(
  temp      = mean(temp,na.rm=T),
  rh        = mean(rh,na.rm=T), 
  wind.sp   = mean(wind.sp,na.rm=T), 
  vis.dist  = mean(vis.dist,na.rm=T), 
  dew.point = mean(dew.point,na.rm=T), 
  lat       = mean(lat,na.rm=T), 
  lon       = mean(lon,na.rm=T), 
  elev      = mean(elev,na.rm=T)
), by = "USAFID"]
```

Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees

```{r}
met_avg[, region := fifelse(lon >= -98 & lat > 39.71, "NE",
                fifelse(lon < -98 & lat > 39.71, "NW",
                fifelse(lon < -98 & lat <= 39.71, "SW","SE")))
    ]
table(met_avg$region)
```


Create a categorical variable for elevation as in the lecture slides

```{r}
met_avg[,elev_cat := fifelse(elev > 252, "high", "low")]
```

## Step3. Make Violin plots of dew point temp by region

```{r violin-dewpoint}
met_avg[!is.na(region)] %>%
  ggplot()+
  geom_violin(mapping = aes(x=1,y=dew.point,color = region,fill = region))+facet_wrap(~region,nrow = 1)
```
From the plot, it can be seen that the highest dew point temperatures are reported in the southeast.

```{r violin-wind.sp}
met_avg[!is.na(region) & !is.na(wind.sp)] %>%
  ggplot()+
  geom_violin(mapping = aes(x=1,y=wind.sp,color = region,fill = region)) +
  facet_wrap(~region, nrow =2)
```

Winds in the southeast and northeest are concentrated at 0 to 2.5



## Step4. Use geom_point with geom_smooth to examine the association between dew point temperature and wind speed by region

Colour points by region
Make sure to deal with NA category
Fit a linear regression line by region
Describe what you observe in the graph

```{r scatterplot-dewpoint-wind.sp}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp, y = dew.point)) + 
  geom_point(mapping = aes(color = region)) + 
  geom_smooth(method = lm, mapping = aes(linetype = region)) +
  facet_wrap(~ region, nrow = 2)

```
the number of dew.point in NE, NW, SE were going up with the increase of wind.sp
the number of dew.point in SW was going down with the increase of wind.sp


## Step5.Use geom_bar to create barplots of the weather stations by elevation category coloured by region


```{r geom_bar}
met_avg[!is.na(region) & !is.na(elev_cat)] %>%
  ggplot() +
  geom_bar(mapping = aes(x = elev_cat, fill = region), position = "dodge") +
  labs(x = "Elevation Category" , y = "Count of Elevation Category ", title = "Elevation Category by Region") +
  scale_fill_brewer(palette = "Purples")
```

```{r barplot- elev_cat}
met_avg[!is.na(region) & !is.na(elev_cat)] %>% 
  ggplot() + 
  labs(title = "Elev_Cat",x = "Elevation Category" , y = "Count of Elevation Category ")+
  geom_bar(mapping = aes(x = elev_cat, fill = region,color= region), position = "dodge")+
  scale_fill_brewer(palette = "Reds") +
  facet_wrap(~elev_cat)
```


About 500 places in northeast region have low elevation
Few places in northwest region have low elevation


## Step6. Use stat_summary to examine mean dew point and wind speed by region with standard deviation error bars

```{r}
met_avg[!is.na(dew.point)] %>%
  ggplot(mapping = aes(x=region,y=dew.point))+
  stat_summary(fun.data = mean_sdl,geom = "pointrange")+
  stat_summary(fun.data = mean_sdl,geom = "errorbar")
```

mean_sdl calculates the mean plus or minus the constant times the standard deviation


The range of dew.point values in the northwest and southwest direction have relatively larger standard deviation and lower mean than others.

```{r}
met_avg[!is.na(wind.sp)] %>%
  ggplot(mapping = aes(x=region,y=wind.sp))+
  stat_summary(fun.data = mean_sdl,geom = "pointrange")+
  stat_summary(fun.data = mean_sdl,geom = "errorbar")
```
##Step7. Generate a map of weather stations and show trend in relative humidity of top 10


```{r}
library(leaflet)
# Generating a color palette
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal

```

```{r}
#Use addMarkers to include the top 10 places in relative h (hint: this will be useful rank(-rh) <= 10)
top10rh <- met_avg[ rank(-rh) <= 10]
```

Here I can use top10rh data subset to show just the weather stations that registered the top 10 relative humidity values in the 1st week of August. For now I’ll show all:



```{r}
rhmap <- leaflet(met_avg) %>% 
  # The looks of the Map
  addProviderTiles('CartoDB.Positron') %>% 
  # Some circles
  addCircles(
    lat = ~lat, lng=~lon,
                                                  # HERE IS OUR PAL!
    label = ~paste0(rh), color = ~ rh.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Relative Humid.', opacity=1)
rhmap
```

 Places with the Top10 highest relative humid are  mainly located on the East side
 
## Step8. 8. Use a ggplot extension 

In ggstance, you supply aesthetics in their natural order:
 
```{r}
if (!require("ggstance")) install.packages("ggstance")
devtools::install_github("lionel-/ggstance")
library("ggstance")
```
 
```{r}
# Horizontal with ggstance
ggplot(met_avg, aes(wind.sp, elev_cat, fill = factor(region))) +
  geom_boxploth()
```






